Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-06-26 PandaEvolution.tex
Negative space added in PandaEvolution
Fix spaces
List of comments to address in each section

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
    \vspace{-0.3in}
    \caption{PanDA Broker architecture as deployed on Titan. Numbers indicates
    the execution process of a detector simulation job, part of the production
    ATLAS MC workflow.}
\label{fig:panda_broker}
\end{figure}

>>>>

Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 2: The authors spent quite a few paragraphs for the detail
	design and implementation of the PanDA WMS, but introduced few background
	of the LHC applications. According to the paper, I understand the
	applications running on Titan are a large amount of small independent
	jobs. Thus, it seems less cost-effective to run these jobs on a
	world-leading supercomputer like Titan which has a fast and expensive
	interconnection. It will be more interesting if the authors can compare
	the solution on Titan (proposed in this paper) with the original
	distributed computing environment and HPECloud from Google mentioned in
	the related work section.
	\item REVIEWER 5: In Section 3.2, GPUs are not considered in this work,
	which would be a waste for using HPC computing. Could this work be
	extended to include GPUs or other heterogenous resources?
\end{enumerate}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/panda_broker_architecture.pdf}
    \vspace{-0.3in}
    \caption{PanDA Broker architecture as deployed on Titan. Numbers
    indicates the execution process of a detector simulation job, part of the
    production ATLAS MC workflow.}
\label{fig:panda_broker}
\end{figure}
\vspace{-0.04in}

### 2017-06-26 conclusion.tex
List of comments to address in each section

% PanDA WMS was developed to meet the scale and complexity of LHC
% distributed computing for the ATLAS experiment. In the process, the old batch
% job paradigm of computing in HEP was discarded in favor of a far more
% flexible and scalable model. The success of PanDA at the LHC is leading to
% widespread adoption and testing by other experiments. PanDA is the first
% exascale workload management system in HEP, already operating at a million
% computing jobs per day, and processing over an exabyte of data in 2013. Next
% LHC run will pose massive computing challenges. With a doubling of the beam
% energy and luminosity as well as an increased need for simulates data, the
% data volume is expected to increase with a factor 5--6 or more. Storing and
% processing this amount of data is a challenge that cannot be resolved with the
% currently existing computing resources in ATLAS\@. To resolve this challenge,
% ATLAS is turning to commercial as well as academic Cloud services and HPCs via
% the PanDA system. Also the work underway is enabling the use of PanDA by new
% scientific collaborations and communities as a means of leveraging extreme
% scale computing resources with a low barrier of entry. The technology base
% provided by the PanDA system will enhance the usage of a variety of
% high-performance computing resources available to basic research.

@@

The 52M core hours used by ATLAS, via PanDA, is over 2\% of the total
utilization on Titan over the same period, bringing the time-averaged
utilization of Titan to be consistently upwards of 90\%. Given that the
average average utilization of most other leadership class machines is less
(e.g., NSF's flagship Blue Waters the average utilization fluctuates between
60-80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
approaches elsewhere. These unprecedented efficiency gains aside, this work is
just a starting point towards more effective operational models for future
leadership and online analytical platforms~\cite{foap-url}. These platforms
will have to support ever increasing complex workloads with varying models for
dynamic resource federation.

% Recent efforts with Google Cloud for the CMS
% experiment -- HEPCloud~\cite{hepcloud,googlehep} represent similar yet
% different approaches to dynamic resource federation. As the type,
% heterogeneity and complexity of future platforms  (e.g., Exascale, Commercial
% Clouds) for science increases, there will be a greater need and emphasis on
% abstractions to ensure WMS can both utilize as well as support evolving
% platforms. Our experience demonstrated the advantages arising from an
% abstractions based approach \textendash{} in particular scalable implementations
% of pilot abstraction.

>>>>

Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 3: How extensible is this? Do you have a sense of backfill
	availability on other systems? Is this unique to Titan or Cray or more
	ubiquitous of an option?
\end{enumerate}

@@

The 52M core hours used by ATLAS, via PanDA, is over 2\% of the total
utilization on Titan over the same period, bringing the time-averaged
utilization of Titan to be consistently upwards of 90\%. Given that the
average average utilization of most other leadership class machines is less
(e.g., NSF's flagship Blue Waters the average utilization fluctuates between
60--80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
approaches elsewhere. These unprecedented efficiency gains aside, this work
is just a starting point towards more effective operational models for future
leadership and online analytical platforms~\cite{foap-url}. These platforms
will have to support ever increasing complex workloads with varying models
for dynamic resource federation.

% Recent efforts with Google Cloud for the CMS experiment --
% HEPCloud~\cite{hepcloud,googlehep} represent similar yet different
% approaches to dynamic resource federation. As the type, heterogeneity and
% complexity of future platforms  (e.g., Exascale, Commercial Clouds) for
% science increases, there will be a greater need and emphasis on
% abstractions to ensure WMS can both utilize as well as support evolving
% platforms. Our experience demonstrated the advantages arising from an
% abstractions based approach \textendash{} in particular scalable
% implementations of pilot abstraction.

### 2017-06-26 introduction.tex
List of comments to address in each section

@@

% . Consequently, Titan now analyzes up to 3.6\% of the simulation workload of
% the ATLAS experiment. The case study presents the design and integration of
% PanDA

>>>>

Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 1: listing the key observations in one place, and in a
	consistent fashion. I found The 5 points listed in the first part of the
	rebuttal quite compelling, and strongly urge the authors to put them in
	the paper in a prominent location.
	\item REVIEWER 1: A glossary would be welcomed.
	\item REVIEWER 5: I would have preferred to see a more detailed list of
	contributions, including the novelty of the proposed approach as compared
	to prior work.
	\item I am ambivalent about the use of the term ``converging'' in the
	paper's title. While the paper does port workloads to high-performance
	computing resources, it does not deal with compute-oriented optimizations
	(parallelization), lowering communication, etc., which are often major
	concerns in HPC programming. In this sense, the paper is closer in spirit
	to how to port high-throughput apps to high-performance clusters. Agreed?
\end{enumerate}

@@

% . Consequently, Titan now analyzes up to 3.6\% of the simulation workload
% of the ATLAS experiment. The case study presents the design and integration
% of PanDA

### 2017-06-26 nge.tex
List of comments to address in each section

>>>>

Comments to address in this section:
\begin{enumerate}
	\item REVIEWER 2: In the weak scalability evaluations (in Figure 7 and
	8), the overhead of the Next Generation Executor (NGE) has grown linearly
	with the number of the nodes. Can you explain why the average time
	execution time of AthenaMP is decreased even when the overhead is
	increasing?
\end{enumerate}

### 2017-06-26 overall.tex
List of comments to address in each section

Currently, ATLAS deploys 20 instances of the PanDA Broker on 4 Titan's DTNs, 5
instances per DTN. Each broker submits and manages the execution of 15 to 300
ATLAS jobs, one job for each Titan's worker node, and a theoretical maximum
concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have operated
only in backfill mode, without a defined time allocation, and running at the
lowest priority on Titan.

We evaluate the efficiency, scalability and reliability of the deployment of
PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and
AthenaMP. We discuss challenges and limitations of our approach at multiple
levels arising from the specifics of workload, middleware and methods. All the
measurements were performed between January 2016 and February 2017, hereafter
called ‘experiment time window

% -----------------------------------------------------------------------------
\subsection{Characterizing the PanDA Broker on Titan}
\label{ssec:broker_titan}

@@

% Usually, detector simulations performed on the Grid process 10 times the
% number of events than those on Titan. This explains the difference between the
% percentage of detector simulations and of events computed by those simulations
% on Titan.
PanDA Broker could fit the number of events to the walltime of each available
backfill slot on the base of the distributions of the time taken by one event to
be simulated. That specific number of event could then be pulled from the PanDA
Event service~\cite{calafiura2015atlas} and given as input to one or more
simulations. Once packaged into the MPI script submitted to titan's PBS batch
system, these simulations would better fit their available backfill slot,
contributing to increase the efficiency of PanDA Brokers.

@@

ATLAS jobs are read heavy: On average, the amount of data read per worker node
is less than 400 MB, while the amount of data written is less than 170 MB.
Distributions of read and written data are different: The read operation
distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
the written amount of data has a very narrow distribution.

>>>>

Comments to address in this section:
\begin{enumerate}
    \item REVIEWER 2: In Figure 4, there are some inconsistencies. The figure
    caption shows there are only 62175 measures, but in the scatter plot
    (right) there are 62555 data points. In addition, according to the
    distribution of the number of work nodes (left), it does not look like
    the mean number of work available is 691 as the figure caption
    illustrates.
    \item REVIEWER 3: t’s interesting that Feb/March have such an increase in
    backfill availability in 2017. The authors attribute this to hardware
    upgrades for the DTNs. It’s mentioned that the load is only 3.6\% now and
    therefore further upgrades will likely not cause similar improvement.
    What was the previous CPU load? Why were particularly low percentages
    utilized in April/May 2016?
    \item REVIEWER 4: “Genat4” at the end of Section 4 is a typo.
    \item REVIEWER 4: Please explain why you used Titan at all, in spite of
    your statement that it is 1/2 to 1/3 as efficient as other resources. Was
    it merely that you were given so much time on this resource that the
    inefficiency doesn’t matter?
    \item REVIEWER 4: Why didn’t you benchmark the performance of your
    workload on other HPC resources that might have delivered better
    performance, prior to reengineering your entire workflow to deal with the
    challenges posed by Titan, a machine that doesn’t seem to justify the
    pain?
    \item REVIEWER 4: The INCITE program operational requirements for the
    LCFs are 90\% uptime. Using large batches of Geant4 jobs that run on a
    single node or less to reach this metric sounds like gaming the system
    rather than an accomplishment.
    \item REVIEWER 5: In Figure 3, it shows the unused core hours are much
    higher than those used by ATLAS. What are the reasons for such low usage?
\end{enumerate}

Currently, ATLAS deploys 20 instances of the PanDA Broker on 4 Titan's DTNs,
5 instances per DTN\@. Each broker submits and manages the execution of 15 to
300 ATLAS jobs, one job for each Titan's worker node, and a theoretical
maximum concurrent use of 96,000 cores. Since November 2015, PanDA Brokers
have operated only in backfill mode, without a defined time allocation, and
running at the lowest priority on Titan.

We evaluate the efficiency, scalability and reliability of the deployment of
PanDA WMS on Titan by characterizing the behavior of both PanDA Broker and
AthenaMP\@. We discuss challenges and limitations of our approach at multiple
levels arising from the specifics of workload, middleware and methods. All
the measurements were performed between January 2016 and February 2017,
hereafter called ‘experiment time window

% ---------------------------------------------------------------------------
\subsection{Characterizing the PanDA Broker on Titan}
\label{ssec:broker_titan}

@@

% Usually, detector simulations performed on the Grid process 10 times the
% number of events than those on Titan. This explains the difference between
% the percentage of detector simulations and of events computed by those
% simulations on Titan.
PanDA Broker could fit the number of events to the walltime of each available
backfill slot on the base of the distributions of the time taken by one event
to be simulated. That specific number of event could then be pulled from the
PanDA Event service~\cite{calafiura2015atlas} and given as input to one or
more simulations. Once packaged into the MPI script submitted to titan's PBS
batch system, these simulations would better fit their available backfill
slot, contributing to increase the efficiency of PanDA Brokers.

@@

ATLAS jobs are read heavy: On average, the amount of data read per worker
node is less than 400 MB, while the amount of data written is less than 170
MB\@. Distributions of read and written data are different: The read
operation distribution per job shows a long tail, ranging from 12.5 GB to 250
GB, while the written amount of data has a very narrow distribution.
